Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-03 paper/paper.qmd
commit changes to paper.qmd

To give an overview of the outcome variable, divided per candidate, we provide summary statistics of this key variable, highlighting the characteristics of the polling support for each candidate across the battleground states. 
```{r}
#| label: tbl-outcomevar
#| fig-cap: "Summary Statistics for Trump and Harris Polling Data - Presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states. All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
#| echo: false
#| warning: false
#| message: false

@@

### Predictive Variables
#### State
The state variable indicates the U.S. state where the poll was conducted or targeted. In the context of the U.S. presidential election, the outcome is frequently determined by several key states. In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party. To enhance our model and refine our predictions, we focused exclusively on these seven battleground states. The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-state-distribution
#| fig-cap: "Bar plot showing the distribution of observations across key battleground states in the 2024 United States Presidential Election polling data. The plot illustrates the count of polling observations for each state, highlighting states with more frequent polling activity. This visualization aids in understanding the level of engagement and data collection in each battleground state leading up to the election."

# Visualize the distribution of the state variable
ggplot(clean_poll_data, aes(x = state)) +
  geom_bar(fill = "steelblue") +  # Bar plot with specified fill color
  labs(
    title = "Distribution of State Variable in Polling Data",
    x = "State",
    y = "Count of Observations"
  ) +
  theme_minimal() +  # Use a minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the plot title
  )
```
The distribution of the state variable is visualized in @fig-state-distribution and indicates that some states, such as Pennsylvania, North Carolina and Wisconsin have more polling observations compared to others, like Nevada. This variability reflects differences in polling frequency and interest among these battleground states. States with more polling observations provide a larger sample size, potentially leading to more stable and reliable predictions for those states in the model.

#### End Date
Another predictive variable is 'end-date'
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-end-date-distribution
#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election. The histogram shows the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches. This visualization is crucial for understanding how polling data correlates with the election timeline."

# Visualize the distribution of end_date
ggplot(clean_poll_data, aes(x = end_date)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +  # Histogram with specified binwidth
  labs(
    title = "Distribution of Polling End Dates",
    x = "End Date",
    y = "Count of Polls"
  ) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 week") +  # Format date labels and breaks
  theme_minimal() +  # Use a minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the plot title
  )

@@

'numeric grade'
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election. Displays the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches. This visualization is crucial for understanding how polling data correlates with the election timeline."

@@

'transparency score'
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-transparency-score-distribution
#| fig-cap: "Histogram illustrating the distribution of transparency scores for polls in key battleground states. Reveals the frequency of different transparency ratings, providing insight into the reliability of polling data. Understanding the distribution of transparency scores is crucial for evaluating the credibility of the polling results as Election Day approaches."

@@

'pollster'
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-pollster-distribution
#| fig-cap: "Horizontal bar graph illustrating the distribution of polls by various pollsters in key battleground states. Displays the count of polls conducted by each pollster, highlighting the most frequently utilized polling organizations. Understanding the diversity of pollsters is essential for assessing the reliability and variance in polling data as Election Day approaches."

# Horizontal bar graph for the distribution of pollsters
ggplot(clean_poll_data, aes(x = factor(pollster))) +  # Convert to factor to ensure proper categorization
  geom_bar(fill = "purple") +
  labs(
       x = "Count of Polls",
       y = "Pollster") +  # Notice the labels are switched for horizontal orientation
  theme_minimal() +
  coord_flip() +  # Flip the coordinates for a horizontal bar graph
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Adjust y-axis text if necessary
```

'start_date'
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-start-date-distribution
#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states in the lead-up to the 2024 United States Presidential Election. This visualization displays the frequency of polls conducted over time, providing insights into when polling activities were initiated and potential trends in voter sentiment as Election Day approaches. Understanding the timing of polls is crucial for evaluating the relevance of the polling data."

>>>>

To give an overview of the outcome variable, divided per candidate, we provide summary statistics of this key variable, highlighting the characteristics of the polling support for each candidate across the battleground states. 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-outcomevar
#| fig-cap: "Summary Statistics for Trump and Harris Polling Data - Presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states. All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."

@@

### Predictive Variables
#### State
The state variable indicates the U.S. state where the poll was conducted or targeted. In the context of the U.S. presidential election, the outcome is frequently determined by several key states. In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party. To enhance our model and refine our predictions, we focused exclusively on these seven battleground states. The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-state-distribution
#| fig-cap: "Bar plot showing the distribution of observations across key battleground states in the 2024 United States Presidential Election polling data. The plot illustrates the count of polling observations for each state, highlighting states with more frequent polling activity. This visualization aids in understanding the level of engagement and data collection in each battleground state leading up to the election."

# Visualize the distribution of the state variable
ggplot(clean_poll_data, aes(x = state)) +
  geom_bar(fill = "steelblue") +  # Bar plot with specified fill color
  labs(
    x = "State",
    y = "Count of Observations"
  ) +
  theme_minimal() +  # Use a minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the plot title
  )
```
The distribution of the state variable is visualized in @fig-state-distribution and indicates that some states, such as Pennsylvania, North Carolina and Wisconsin have more polling observations compared to others, like Nevada. This variability reflects differences in polling frequency and interest among these battleground states. States with more polling observations provide a larger sample size, potentially leading to more stable and reliable predictions for those states in the model.

#### End Date
Another predictive variable is 'end-date'
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-end-date-distribution
#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election. The histogram shows the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches. This visualization is crucial for understanding how polling data correlates with the election timeline."

# Visualize the distribution of end_date
ggplot(clean_poll_data, aes(x = end_date)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +  # Histogram with specified binwidth
  labs(
    x = "End Date",
    y = "Count of Polls"
  ) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 week") +  # Format date labels and breaks
  theme_minimal() +  # Use a minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the plot title
  )

@@

#### Numeric Grade
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election. Displays the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches. This visualization is crucial for understanding how polling data correlates with the election timeline."

@@

#### Transparency Score
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-transparency-score-distribution
#| fig-cap: "Histogram illustrating the distribution of transparency scores for polls in key battleground states. Reveals the frequency of different transparency ratings, providing insight into the reliability of polling data. Understanding the distribution of transparency scores is crucial for evaluating the credibility of the polling results as Election Day approaches."

@@

#### Pollster
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-pollster-distribution
#| fig-cap: "Horizontal bar graph illustrating the distribution of polls by various pollsters in key battleground states. Displays the count of polls conducted by each pollster, highlighting the most frequently utilized polling organizations. Understanding the diversity of pollsters is essential for assessing the reliability and variance in polling data as Election Day approaches."

# Horizontal bar graph for the distribution of pollsters
ggplot(clean_poll_data, aes(x = factor(pollster))) +  # Convert to factor to ensure proper categorization
  geom_bar(fill = "purple") +
  labs(
       x = "Count of Polls",
       y = "Pollster") +  
  theme_minimal() +
  coord_flip() +  
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  
```

#### Start Date
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-start-date-distribution
#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states in the lead-up to the 2024 United States Presidential Election. This visualization displays the frequency of polls conducted over time, providing insights into when polling activities were initiated and potential trends in voter sentiment as Election Day approaches. Understanding the timing of polls is crucial for evaluating the relevance of the polling data."
